---
title: "12. Model Tunning and the Dangers of Overfitting"
format: html
date: last-modified
execute:
    cache: true
code-fold: true
toc: true
---

```{r}
#| label: code

library(tidymodels)
data(ames)
ames <- mutate(ames, Sale_Price = log10(Sale_Price))

set.seed(502)
ames_split <- initial_split(ames, prop = 0.80, strata = Sale_Price)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)

ames_rec <-
  recipe(
    Sale_Price ~ Neighborhood +
      Gr_Liv_Area +
      Year_Built +
      Bldg_Type +
      Latitude +
      Longitude,
    data = ames_train
  ) %>%
  step_log(Gr_Liv_Area, base = 10) %>%
  step_other(Neighborhood, threshold = 0.01) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_interact(~ Gr_Liv_Area:starts_with("Bldg_Type_")) %>%
  step_ns(Latitude, Longitude, deg_free = 20)

lm_model <- linear_reg() %>% set_engine("lm")

lm_wflow <-
  workflow() %>%
  add_model(lm_model) %>%
  add_recipe(ames_rec)

lm_fit <- fit(lm_wflow, ames_train)

rf_model <-
  rand_forest(trees = 1000) %>%
  set_engine("ranger") %>%
  set_mode("regression")

rf_wflow <-
  workflow() %>%
  add_formula(
    Sale_Price ~ Neighborhood +
      Gr_Liv_Area +
      Year_Built +
      Bldg_Type +
      Latitude +
      Longitude
  ) %>%
  add_model(rf_model)

set.seed(1001)
ames_folds <- vfold_cv(ames_train, v = 10)

keep_pred <- control_resamples(save_pred = TRUE, save_workflow = TRUE)

set.seed(1003)
rf_res <- rf_wflow %>%
  fit_resamples(resamples = ames_folds, control = keep_pred)
```

## 12. Model Tuning and the Dangers of Overfitting


```{r}
#| label: optimize-link-function
#| eval: false

llhood <- function(...) {
  logistic_reg() %>%
    set_engine("glm", ...) %>%
    fit(Class ~ ., data = training_set) %>%
    glance() %>%
    select(logLik)
}

bind_rows(
  llhood(),
  llhood(family = binomial(link = "probit")),
  llhood(family = binomial(link = "cloglog"))
) %>%
  mutate(link = c("logit", "probit", "c-log-log")) %>%
  arrange(desc(logLik))

#> # A tibble: 3 × 2
#>   logLik link
#>    <dbl> <chr>
#> 1  -258. logit
#> 2  -262. probit
#> 3  -270. c-log-log
```

Scale of values makes it difficult to know if these differences are important. Can improve my resampling.

```{r}
#| label: optimize-link-function-resample
#| eval: false

set.seed(1201)
rs <- vfold_cv(training_set, repeats = 10)

# Return the individual resampled performance estimates:
lloss <- function(...) {
  perf_meas <- metric_set(roc_auc, mn_log_loss)

  logistic_reg() %>%
    set_engine("glm", ...) %>%
    fit_resamples(Class ~ A + B, rs, metrics = perf_meas) %>%
    collect_metrics(summarize = FALSE) %>%
    select(id, id2, .metric, .estimate)
}

resampled_res <-
  bind_rows(
    lloss() %>% mutate(model = "logistic"),
    lloss(family = binomial(link = "probit")) %>% mutate(model = "probit"),
    lloss(family = binomial(link = "cloglog")) %>% mutate(model = "c-log-log")
  ) %>%
  # Convert log-loss to log-likelihood:
  mutate(
    .estimate = ifelse(.metric == "mn_log_loss", -.estimate, .estimate)
  ) %>%
  group_by(model, .metric) %>%
  summarize(
    mean = mean(.estimate, na.rm = TRUE),
    std_err = sd(.estimate, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

resampled_res %>%
  filter(.metric == "mn_log_loss") %>%
  ggplot(aes(x = mean, y = model)) +
  geom_point() +
  geom_errorbar(
    aes(xmin = mean - 1.64 * std_err, xmax = mean + 1.64 * std_err),
    width = .1
  ) +
  labs(y = NULL, x = "log-likelihood")
```

```{r}
#| label: tuning

ames_rec <-
  recipe(
    Sale_Price ~ Neighborhood +
      Gr_Liv_Area +
      Year_Built +
      Bldg_Type +
      Latitude +
      Longitude,
    data = ames_train
  ) %>%
  step_log(Gr_Liv_Area, base = 10) %>%
  step_other(Neighborhood, threshold = tune()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_interact(~ Gr_Liv_Area:starts_with("Bldg_Type_")) %>%
  step_ns(Longitude, deg_free = tune("longitude df")) %>%
  step_ns(Latitude, deg_free = tune("latitude df"))

recipes_param <- extract_parameter_set_dials(ames_rec)
recipes_param
```

Can call a step more than once to optimize different preds differently. E.g., **step_ns()**  is called twice to tune lat and long to different levels of smoothness. **tune()** can take a name argument to distinguish between multiple calls of the same step. 

```{r}
#| label: tuning-nondefaulting-params

rf_spec <-
  rand_forest(mtry = tune()) %>%
  set_engine("ranger", regularization.factor = tune("regularization")) %>%
  set_mode("regression")

rf_param <- extract_parameter_set_dials(rf_spec)

pca_rec <-
  recipe(Sale_Price ~ ., data = ames_train) %>%
  # Select the square-footage predictors and extract their PCA components:
  step_normalize(contains("SF")) %>%
  # Select the number of components needed to capture 95% of
  # the variance in the predictors.
  step_pca(contains("SF"), threshold = .95)

updated_param <-
  workflow() %>%
  add_model(rf_spec) %>%
  add_recipe(pca_rec) %>%
  extract_parameter_set_dials() %>%
  finalize(ames_train)
updated_param

updated_param %>% extract_parameter_dials("mtry")
```

Use **finalize()** to tune params whose ranges can change during the workflow, like when using PCA to create new preds.