---
title: "13. Grid Search"
format: html
date: last-modified
execute:
    cache: true
code-fold: true
toc: true
---

```{r}
#| label: code

library(tidymodels)
data(ames)
ames <- mutate(ames, Sale_Price = log10(Sale_Price))

set.seed(502)
ames_split <- initial_split(ames, prop = 0.80, strata = Sale_Price)
ames_train <- training(ames_split)
ames_test <- testing(ames_split)

ames_rec <-
  recipe(
    Sale_Price ~ Neighborhood +
      Gr_Liv_Area +
      Year_Built +
      Bldg_Type +
      Latitude +
      Longitude,
    data = ames_train
  ) %>%
  step_log(Gr_Liv_Area, base = 10) %>%
  step_other(Neighborhood, threshold = 0.01) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_interact(~ Gr_Liv_Area:starts_with("Bldg_Type_")) %>%
  step_ns(Latitude, Longitude, deg_free = 20)

lm_model <- linear_reg() %>% set_engine("lm")

lm_wflow <-
  workflow() %>%
  add_model(lm_model) %>%
  add_recipe(ames_rec)

lm_fit <- fit(lm_wflow, ames_train)

rf_model <-
  rand_forest(trees = 1000) %>%
  set_engine("ranger") %>%
  set_mode("regression")

rf_wflow <-
  workflow() %>%
  add_formula(
    Sale_Price ~ Neighborhood +
      Gr_Liv_Area +
      Year_Built +
      Bldg_Type +
      Latitude +
      Longitude
  ) %>%
  add_model(rf_model)

set.seed(1001)
ames_folds <- vfold_cv(ames_train, v = 10)

keep_pred <- control_resamples(save_pred = TRUE, save_workflow = TRUE)

set.seed(1003)
rf_res <- rf_wflow %>%
  fit_resamples(resamples = ames_folds, control = keep_pred)
```

## 13. Grid Search

Using a multi-layer perceptron (MLP) model (AKA single layer artificial NN) as an example, this section demonstrates how to set up a grid search for hyperparameter tuning using the `tidymodels` framework in R.

```{r}
#| label: setup

mlp_spec <- 
  mlp(hidden_units = tune(), penalty = tune(), epochs = tune()) %>% 
  set_engine("nnet", trace = 0) %>% 
  set_mode("classification")
```

Checking default parameter values for the MLP model specification.

```{r}
#| label: extract-param-defaults

mlp_param <- extract_parameter_set_dials(mlp_spec)

mlp_param %>% extract_parameter_dials("hidden_units")

mlp_param %>% extract_parameter_dials("penalty")

mlp_param %>% extract_parameter_dials("epochs")
```

Regular grid search over three hyperparameters: `hidden_units`, `penalty`, and `epochs`.

```{r}
#| label: reg-grid-crossing

crossing(
  hidden_units = 1:3,
  penalty = c(0.0, 0.1),
  epochs = c(100, 200)
)
```

