---
title: "16. Dimensionality Reduction"
format: html
date: last-modified
execute:
    cache: true
code-fold: true
toc: true
---

## 16.2. Beans

The beans data set contains measurements on different varieties of beans. The goal is to classify the variety based on the measurements. There are many variables trying to describe the shape of bean like area, perimenter, compactness, etc. Many of these variables are correlated. This makes the beans data set a good candidate for dimensionality reduction.

Loading data.

```{r}
#| label: load-data

library(tidymodels)
tidymodels_prefer()
library(beans)
```

Splitting.

```{r}
#| label: data-split

set.seed(1601)
bean_split <- initial_validation_split(
  beans,
  strata = class,
  prop = c(0.75, 0.125)
)

bean_split

bean_train <- training(bean_split)
bean_test <- testing(bean_split)
bean_validation <- validation(bean_split)


set.seed(1602)
bean_val <- validation_set(bean_split)
bean_val$splits[[1]]
```

Before doing dimensionality reduction, let's look at correlations.

```{r}
#| label: bean-corrplot

library(corrplot)
tmwr_cols <- colorRampPalette(c("#91CBD765", "#CA225E"))
bean_train %>%
  select(-class) %>%
  cor() %>%
  corrplot(col = tmwr_cols(200), tl.col = "black", method = "ellipse")
```

## 16.3. A Starter Recipe

Skew is bad for variance calculations like PCA. Using **bestNormalize** package to force symmetric distributions.

```{r}
#| label: bean-recipe

library(bestNormalize)

bean_rec <-
  # Use the training data from the bean_val split object
  recipe(class ~ ., data = bean_train) %>%
  step_zv(all_numeric_predictors()) %>%
  step_orderNorm(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors())
```

## 16.4. Recipes in the Wild

*prep()* is similar to *fit()*, but for recipes.

```{r}
#| label: bean-recipe-train

bean_rec_trained <- prep(bean_rec)
bean_rec_trained
```

Can use *verbose* and/or *log_changes* args to help troubleshoot.

```{r}
#| label: troubleshoot-recipe

bean_rec_trained %>%
  step_dummy(cornbread) %>% # <- not a real predictor
  prep(verbose = TRUE)

show_variables <-
  bean_rec %>%
  prep(log_changes = TRUE)
```

Can use *bake()* to apply the recipe to new data.

```{r}
#| label: bake-recipe

bean_val_processed <- bake(bean_rec_trained, new_data = bean_validation)
```

Histograms before and after recipe.

```{r}
#| label: histograms-recipe

library(patchwork)
p1 <-
  bean_validation %>%
  ggplot(aes(x = area)) +
  geom_histogram(bins = 30, color = "white", fill = "blue", alpha = 1 / 3) +
  ggtitle("Original validation set data")

p2 <-
  bean_val_processed %>%
  ggplot(aes(x = area)) +
  geom_histogram(bins = 30, color = "white", fill = "red", alpha = 1 / 3) +
  ggtitle("Processed validation set data")

p1 + p2
```

## 16.5. Feature Extraction Techniques

```{r}
#| label: FUNC-plot-validation-results

library(ggforce)

plot_validation_results <- function(recipe, dat = bean_validation) {
  recipe %>%
    # Estimate any additional steps
    prep() %>%
    # Process the data (the validation set by default)
    bake(new_data = dat) %>%
    # Create the scatterplot matrix
    ggplot(aes(x = .panel_x, y = .panel_y, color = class, fill = class)) +
    geom_point(alpha = 0.4, size = 0.5) +
    geom_autodensity(alpha = .3) +
    facet_matrix(vars(-class), layer.diag = 2) +
    scale_color_brewer(palette = "Dark2") +
    scale_fill_brewer(palette = "Dark2")
}
```

### 16.5.1. Principal Component Analysis (PCA)

PCA tries to account for as much variance as possible in the first few components.

```{r}
#| label: pca-beans

bean_rec_trained %>%
  step_pca(all_numeric_predictors(), num_comp = 4) %>%
  plot_validation_results() +
  ggtitle("Principal Component Analysis")
```

PC1 and PC2 separate the classes pretty well. Might expect the overal problem of classifying bean variety to be relatively easy.

```{r}
#| label: pca-loadings

library(learntidymodels)

bean_rec_trained %>%
  step_pca(all_numeric_predictors(), num_comp = 4) %>%
  prep() %>%
  plot_top_loadings(component_number <= 4, n = 5) +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Principal Component Analysis")
```


### 16.5.2. Partial Least Squares (PLS)

PLS is a supervised dimensionality reduction technique. It uses the outcome variable to help find components that are good at predicting the outcome.

```{r}
#| label: pls-beans

bean_rec_trained %>%
  step_pls(all_numeric_predictors(), outcome = "class", num_comp = 4) %>%
  plot_validation_results() +
  ggtitle("Partial Least Squares")
```

PLS1 and PLS2 are very similar to PCA1 and PCA2.

```{r}
#| label: pls-loadings

bean_rec_trained %>%
  step_pls(all_numeric_predictors(), outcome = "class", num_comp = 4) %>%
  prep() %>%
  plot_top_loadings(component_number <= 4, n = 5, type = "pls") +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Partial Least Squares")
```

## 16.5.3. Independent Component Analysis (ICA)

ICA finds components that are statistically independent, instead of uncorrelated like PCA.

```{r}
#| label: ica-beans

bean_rec_trained %>%
  step_ica(all_numeric_predictors(), num_comp = 4) %>%
  plot_validation_results() +
  ggtitle("Independent Component Analysis")
```

Not much separation of classes. These independent components do not separate bean types.

### 16.5.4. Uniform Manifold Approximation and Projection (UMAP)

UMAP is a nonlinear dimensionality reduction technique that tries to preserve local structure in the data.

```{r}
#| label: umap-beans

library(embed)

bean_rec_trained %>%
  step_umap(all_numeric_predictors(), num_comp = 4) %>%
  plot_validation_results() +
  ggtitle("UMAP")
```

Between-cluster space is pronounced, but there's some heterogeneous mixture of classes.

Also a supervised version of UMAP:

```{r}
#| label: supervised-umap-beans

bean_rec_trained %>%
  step_umap(all_numeric_predictors(), outcome = "class", num_comp = 4) %>%
  plot_validation_results() +
  ggtitle("UMAP (supervised)")
```

Looks promising.

UMAP is powerful but very sensitive to tuning params (e.g., number of neighbors, min dist). The defaults may not be optimal.

## 16.6. Modeling

Trying out dimensionality reduction techniques in models.

```{r}
#| label: modeling-beans

library(baguette)
library(discrim)

mlp_spec <-
  mlp(hidden_units = tune(), penalty = tune(), epochs = tune()) %>%
  set_engine('nnet') %>%
  set_mode('classification')

bagging_spec <-
  bag_tree() %>%
  set_engine('rpart') %>%
  set_mode('classification')

fda_spec <-
  discrim_flexible(
    prod_degree = tune()
  ) %>%
  set_engine('earth')

rda_spec <-
  discrim_regularized(frac_common_cov = tune(), frac_identity = tune()) %>%
  set_engine('klaR')

bayes_spec <-
  naive_Bayes() %>%
  set_engine('klaR')
```

Needs recipes for the dim reduction techniques we're trying.

```{r}
#| label: modeling-recipes

bean_rec <-
  recipe(class ~ ., data = bean_train) %>%
  step_zv(all_numeric_predictors()) %>%
  step_orderNorm(all_numeric_predictors()) %>%
  step_normalize(all_numeric_predictors())

pls_rec <-
  bean_rec %>%
  step_pls(all_numeric_predictors(), outcome = "class", num_comp = tune())

umap_rec <-
  bean_rec %>%
  step_umap(
    all_numeric_predictors(),
    outcome = "class",
    num_comp = tune(),
    neighbors = tune(),
    min_dist = tune()
  )
```

Creating workflow set and tuning:

```{r}
#| label: workflow-set-beans
#| eval: false

# time-consuming code commented out to speed up rendering
# ctrl <- control_grid(parallel_over = "everything")
# bean_res <-
#   workflow_set(
#     preproc = list(basic = class ~., pls = pls_rec, umap = umap_rec),
#     models = list(bayes = bayes_spec, fda = fda_spec,
#                   rda = rda_spec, bag = bagging_spec,
#                   mlp = mlp_spec)
#   ) %>%
#   workflow_map(
#     verbose = TRUE,
#     seed = 1603,
#     resamples = bean_val,
#     grid = 10,
#     metrics = metric_set(roc_auc),
#     control = ctrl
#   )
```

Visualizing results.

```{r}
#| label: visualize-results-beans
#| eval: false

rankings <-
  rank_results(bean_res, select_best = TRUE) %>%
  mutate(method = map_chr(wflow_id, ~ str_split(.x, "_", simplify = TRUE)[1]))

tidymodels_prefer()
filter(rankings, rank <= 5) %>% dplyr::select(rank, mean, model, method)
#> # A tibble: 5 × 4
#>    rank  mean model               method
#>   <int> <dbl> <chr>               <chr>
#> 1     1 0.996 mlp                 pls
#> 2     2 0.996 discrim_regularized pls
#> 3     3 0.995 discrim_flexible    basic
#> 4     4 0.995 naive_Bayes         pls
#> 5     5 0.994 naive_Bayes         basic
```

Results of multiclass ROC AUC on testing set:

```{r}
#| label: eval-metric
#| eval: false

collect_metrics(rda_res)
```